\def\duedate{\today}
\def\HWnum{7}
\input{../preamble.tex}

\begin{document}

\prob{1 -- Chapter 6 \# 1}{

Let $\ket{\phi_{n}}$ be the eigenstates of a Hamiltonian $\hat{H}$ (a hermitian operator).
Assume that the $\ket{\phi_{n}}$ form a discrete orthonormal basis.
Consider the operator $\hat{U}(m,n)$ defined as
\begin{eqnarray}
    \hat{U}(m,n) = \ket{\phi_{m}} \bra{\phi_{n}}
.\end{eqnarray}

(a) Obtain the adjoint $\hat{U}^{\dagger}(m,n)$.

(b) Evaluate the commutator $[ \hat{H}, \hat{U}(m,n) ]$.

(c) Show that
\begin{eqnarray}
    \hat{U}(m,n) \hat{U}^{\dagger}(p,q) = \delta_{nq} \hat{U}(m,p)
.\end{eqnarray}

(d) Calculate the trace of $\hat{U}(m,n)$.
The trace of an operator $\hat{A}$ is defined as
\begin{eqnarray}
    \Tr(\hat{A}) = \sum_{n} \bra{\phi_{n}} \hat{A} \ket{\phi_{n}}
.\end{eqnarray}

(e) If $\hat{A}$ is an operator with matrix elements $A_{mn} = \bra{\phi_{m}} \hat{A} \ket{\phi_{n}}$, show that
\begin{eqnarray}
    \hat{A} = \sum_{m,n} A_{mn} \hat{U}(m,n)
.\end{eqnarray}

(f) Show that 
\begin{eqnarray}
    A_{pq} = \Tr[ \hat{A} \hat{U}^{\dagger}(p,q) ]
.\end{eqnarray}

}

\sol{

(a) The adjoint $A^{\dagger}$ is defined such that $\bra{\psi} A^{\dagger}$ is dual to $A \ket{\psi}$.
Notice that $\hat{U}(m,n) \ket{\psi} = \ket{\phi_{m}} \bra{\phi_{n}}\ket{\psi}$ whose dual is $\bra{\psi} \hat{U}^{\dagger} = \bra{\phi_{n}}\ket{\psi}^{*} \bra{\phi_{m}} = \bra{\psi} \ket{\phi_{n}}\bra{\phi_{m}}$.
Thus, $\hat{U}^{\dagger}(m,n) = \ket{\phi_{n}} \bra{\phi_{m}} = \hat{U}(n,m)$.

(b) The commutator
\begin{eqnarray}
\begin{aligned}
    [\hat{H},\hat{U}(m,n)] &= \hat{H} \ket{\phi_{m}}\bra{\phi_{n}} - \ket{\phi_{m}}\bra{\phi_{n}} \hat{H} = E_{m} \ket{\phi_{m}}\bra{\phi_{n}} - \ket{\phi_{m}}\bra{\phi_{n}} E_{n} \\
                           &= (E_{m} - E_{n}) \hat{U}(m,n)
.\end{aligned}
\end{eqnarray}

(c) Using the fact that $\{ \ket{\phi_{n}} \} $ is orthonormal
\begin{eqnarray}
    \hat{U}(m,n) \hat{U}^{\dagger}(p,q) = \ket{\phi_{m}}\bra{\phi_{n}} \ket{\phi_{q}} \bra{\phi_{p}} = \delta_{nq} \ket{\phi_{m}} \bra{\phi_{p}} = \delta_{nq} \hat{U}(m,p)
.\end{eqnarray}

(d) The trace of $\hat{U}(m,n)$ is 
\begin{eqnarray}
    \Tr(\hat{U}(m,n)) = \sum_{k} \bra{\phi_{k}} \ket{\phi_{m}}\bra{\phi_{n}} \ket{\phi_{k}} = \sum_{k} \delta_{km} \delta_{nk} = \delta_{nm}
.\end{eqnarray}

(e) Since $\{ \ket{\phi_{n}} \} $ forms an orthonormal basis
\begin{eqnarray}
    \hat{A} = \hat{\id} \hat{A} \hat{\id} = \sum_{n,m} \ket{\phi_{m}} \underbrace{\bra{\phi_{m}} \hat{A} \ket{\phi_{n}}}_{A_{mn}} \bra{\phi_{n}} = \sum_{n,m} A_{mn} \hat{U}(m,n)
.\end{eqnarray}

(f) Using the results above, we find
\begin{eqnarray}
    \hat{A} \hat{U}^{\dagger}(p,q) = \sum_{m,n} A_{mn} \hat{U}(m,n) \hat{U}^{\dagger}(p,q) = \sum_{m} A_{mq} \hat{U}(m,p)
\end{eqnarray}
and
\begin{eqnarray}
    \Tr(\hat{A} \hat{U}^{\dagger}(p,q)) = \sum_{m} A_{mq} \delta_{mp} = A_{pq}
.\end{eqnarray}

}


\prob{2 -- Chapter 7 \# 4}{

Consider the Hamiltonian $\hat{H}$ of a particle in a one-dimensional problem given by
\begin{eqnarray}
    \hat{H} = \frac{\hat{p}^2}{2m} + \hat{V}(\hat{x})
,\end{eqnarray}
where $\hat{x}$ and $\hat{p}$ are the position and momentum operators satisfying the standard commutation relations.
Let $\ket{\phi_{n}}$ be the eigenstates of $\hat{H}$ with $\hat{H} \ket{\phi_{n}} = E_{n} \ket{\phi_{n}}$, where $n$ is a discrete index.

(a) By considering the commutator $[\hat{x},\hat{H}]$, show that 
\begin{eqnarray}
    \bra{\phi_{m}} \hat{p} \ket{\phi_{n}} = \alpha_{mn} \bra{\phi_{m}} \hat{x} \ket{\phi_{n}}
,\end{eqnarray}
where the coefficient $\alpha_{mn}$ depends on the energy difference $E_{m} - E_{n}$.

(b) Using the closure relation satisfied by the eigenstates of $\hat{H}$ and the result above, deduce the following relation (sum rule)
\begin{eqnarray}
    \sum_{n} (E_{m} - E_{n})^2 | \bra{\phi_{m}} \hat{x} \ket{\phi_{n}} |^2 = \frac{\hbar^2}{m^2} \bra{\phi_{m}} \hat{p}^2 \ket{\phi_{m}}
.\end{eqnarray}


}

\sol{

(a) Consider the generic operators $\hat{A}, \hat{B}, \hat{C}$.
The commutator
\begin{eqnarray}
    \begin{aligned}
        [\hat{A},\hat{B} \hat{C}] &= \hat{A}\hat{B}\hat{C} - \hat{B}\hat{C}\hat{A} \\
                                  &= \hat{A}\hat{B}\hat{C} - \hat{B}\hat{A}\hat{C} + \hat{B}\hat{A}\hat{C} - \hat{B}\hat{C}\hat{A} \\
                                  &= [\hat{A},\hat{B}]\hat{C} + \hat{B}[\hat{A},\hat{C}]
    .\end{aligned}
\end{eqnarray}
We can use this and the assumption that $[\hat{x},\hat{p}] = i\hbar$ to find
\begin{eqnarray}
    [\hat{x},\hat{p}^2] = [\hat{x},\hat{p}] \hat{p}  + \hat{p}[\hat{x},\hat{p}] = 2i\hbar \hat{p}
,\end{eqnarray}
and therefore,
\begin{eqnarray}
    [\hat{x},\hat{H}] = \frac{1}{2m} [\hat{x},\hat{p}^2] + [\hat{x},\hat{V}(\hat{x})] = \frac{i\hbar}{m}\hat{p}
.\end{eqnarray}
Thus, the matrix elements of the commutator
\begin{eqnarray}
    \bra{\phi_{m}} [\hat{x},\hat{H}] \ket{\phi_{n}} = ( E_{n} - E_{m} ) \bra{\phi_{m}} \hat{x} \ket{\phi_{n}} = \frac{i\hbar}{m} \bra{\phi_{m}} \hat{p} \ket{\phi_{n}}
.\end{eqnarray}
Rearranging, we find
\begin{eqnarray}
    \bra{\phi_{m}} \hat{p} \ket{\phi_{n}} = \underbrace{ \frac{i m}{\hbar} (E_{m} - E_{n}) }_{\alpha_{mn}} \bra{\phi_{m}} \hat{x} \ket{\phi_{n}}
.\end{eqnarray}

(b) Observe the following:
\begin{eqnarray}
    (E_{m} - E_{n})^2 |\bra{\phi_{m}} \hat{x} \ket{\phi_{n}}|^2 = \frac{\hbar^2}{m^2}|\bra{\phi_{m}} \hat{p} \ket{\phi_{n}}|^2 = \frac{\hbar^2}{m^2} \bra{\phi_{m}} \hat{p} \ket{\phi_{n}} \bra{\phi_{n}} \hat{p} \ket{\phi_{m}}
.\end{eqnarray}
If we sum over $n$ on both sides we have
\begin{eqnarray}
    \sum_{n} (E_{m} - E_{n})^2 | \bra{\phi_{m}} \hat{x} \ket{\phi_{n}} |^2 = \frac{\hbar^2}{m^2} \bra{\phi_{m}} \hat{p}^2 \ket{\phi_{m}}
,\end{eqnarray}
where we have used the fact that the states $\{ \ket{\phi_{n}} \} $ are complete on the right.
Note that this is not possible on the l.h.s. since the factor of the energy difference depends on $n$.

}


\prob{3 -- Chapter 6 \# 13}{

Consider a three-dimensional state space.
If a certain set of orthonormal kets $\ket{\phi_{1}}$, $\ket{\phi_2}$, and $\ket{\phi_3}$ are used as the base kets, the operators $\hat{A}$ and $\hat{B}$ are represented by
\begin{eqnarray}
    A = \begin{pmatrix}
        a & 0 & 0 \\
        0 & -a & 0 \\
        0 & 0 & -a
    \end{pmatrix}
    , \quad
    B = \begin{pmatrix}
        b & 0 & 0 \\
        0 & 0 & -ib \\
        0 & ib & 0
    \end{pmatrix}
,\end{eqnarray}
where $a$ and $b$ are real.

(a) It is obvious that $\hat{A}$ has a degenerate spectrum.
Is the spectrum of $\hat{B}$ also degenerate?

(b) Show that $\hat{A}$ and $\hat{B}$ commute.

(c) Find a new set of orthonormal kets which are simultaneous eigenstates of both $\hat{A}$ and $\hat{B}$.
Specify the eigevalues of $\hat{A}$ and $\hat{B}$ for each of these three eigenstates.
Does specifying these eigenvalues uniquely identify the relative common eigenstate?
That is, do $\hat{A}$ and $\hat{B}$ form a complete set of commuting observables?

}

\sol{

(a) Let $\ket{\phi_1}$ correspond to eigenvalue $a$ of $\hat{A}$ and $\ket{\phi_2},\ket{\phi_3}$ correspond to $-a$.
Thus, $\ket{\phi_1}$ also corresopnds to eigenvalue $b$ of $\hat{B}$.
The other eigenvalues of $\hat{B}$ will have eigenvectors which are linear combinations of $\ket{\phi_2},\ket{\phi_3}$, meaning that we only have to diagonalize
\begin{eqnarray}
    \begin{pmatrix}
        0 & -ib \\
        ib & 0
    \end{pmatrix}
.\end{eqnarray}
It should be clear that this matrix has characteristic equation $\lambda^2 - b^2 = 0$, which has roots $\lambda = \pm b$.
Thus, $\hat{B}$ also has a degenerate spectrum.

(b) Taking the matrix products explicitly, we find 
\begin{eqnarray}
    AB - BA = \begin{pmatrix}
        ab & 0 & 0 \\
        0 & 0 & iab \\
        0 & -iab & 0
    \end{pmatrix}
    - 
    \begin{pmatrix}
        ba & 0 & 0 \\
        0 & 0 & iba \\
        0 & -iba & 0
    \end{pmatrix}
    = 0
.\end{eqnarray}
Since the mapping between the matrix representation and hilbert space are bijective, the commutation holds for the operators $\hat{A}$ and $\hat{B}$ in the hilbert space.

(c) In part (a), we found the spectrum of $\hat{B}$.
Now, we solve for the eigenvectors in the supspace spanned by $\ket{\phi_2}$ and $\ket{\phi_3}$, which have the general form $\begin{pmatrix} \alpha_1 & \alpha_2 \end{pmatrix}^{T}$:
\begin{eqnarray}
    -ib \alpha_2 = \pm b \alpha_1 \Rightarrow \alpha_2 = \pm i \alpha_1
.\end{eqnarray}
The eigenvectors corresponding to eigenvalues $\pm b$ is then 
\begin{eqnarray}
    \frac{1}{\sqrt{2}} \Big[ \ket{\phi_2} \pm i \ket{\phi_3} \Big]
.\end{eqnarray}
It should be clear that these eigenvectors of $\hat{B}$ are still eigenvectors of $\hat{A}$ corresponding to eigenvalue $-a$ and also that they are orthogonal to each other and $\ket{\phi_1}$.
Furthermore, the correspondence between eigenvalues and eigenvectors is given by
\begin{align}
    \{ a,b \} &\leftrightarrow \ket{\phi_1} \\
    \{ -a, b \}  &\leftrightarrow (\ket{\phi_2} + i \ket{\phi_3})/\sqrt{2} \\
    \{ -a,-b \}  &\leftrightarrow (\ket{\phi_2} - i\ket{\phi_3})/\sqrt{2}
.\end{align}
It is clear that specifying the eigenvalues of $\hat{A}$ and $\hat{B}$ uniquely specifies the simultaneous corresponding eigenvector of $\hat{A}$ and $\hat{B}$, meaning that $\hat{A}$ and $\hat{B}$ form a complete set of commuting observables.


}


\prob{4 -- Chapter 6 \# 14}{

A molecule is composed of six identical atoms $A_1, A_2, \ldots, A_6$ which form a regular hexagon.
Consider an electron which can be localized on each of the atoms.
Denote with $\ket{\psi_{n}}$ the state in which the electron is localized on the $n^{\rm th}$ atom ($n = 1,2,\ldots,6$).
The electron states will be limited to the space spanned by the $\ket{\psi_{n}}$, assumed to be orthonormal $\bra{\psi_{m}}\ket{\psi_{n}} = \delta_{mn}$; in other words, these six states form a basis.

(a) Define the operator $\hat{R}$ by the following relations:
\begin{eqnarray}
    \hat{R} \ket{\psi_1} = \ket{\psi_2}, \quad \hat{R} \ket{\psi_2} = \ket{\psi_3}, \quad \ldots, \quad \hat{R} \ket{\psi_6} = \ket{\psi_1}
.\end{eqnarray}
Find the eigenvalues and eigenstates of $\hat{R}$.
Show that the eigenvectors form an orthonormal set (i.e. they form a basis).

(b) Show that the adjoint operator $\hat{R}^{\dagger}$ gives
\begin{eqnarray}
    \hat{R}^{\dagger} \ket{\psi_1} = \ket{\psi_6}, \quad \hat{R}^{\dagger} \ket{\psi_2} = \ket{\psi_1}, \quad \ldots, \quad \hat{R}^{\dagger} \ket{\psi_6} = \ket{\psi_5}
.\end{eqnarray}
Show that $\hat{R}$ is unitary.

(c) When the probability of the electron jumping from one site to a contiguous one to the left or right is neglected, its energy is described by the Hamiltonian $\hat{H}_{0}$, whose eigenstates are the six states $\ket{\psi_{n}}$, all with the same eigenvalue $E_0$, namely $\hat{H}_{0} \ket{\psi_{n}} = E_0 \ket{\psi_{n}}$.
The possibility for the electron to jump from one site to another is modeled by adding the Hamiltonian $\hat{H}_{0}$ a perturbation $\hat{V}$ such that 
\begin{eqnarray}
    \begin{aligned}
    &\hat{V} \ket{\psi_1} = -a \ket{\psi_6} - a \ket{\psi_2}, \quad \hat{V} \ket{\psi_2} = -a \ket{\psi_1} - a \ket{\psi_3}, \quad \ldots, \\
    &\hat{V} \ket{\psi_6} = -a \ket{\psi_5} - a \ket{\psi_1}
    .\end{aligned}
\end{eqnarray}
Show that $\hat{R}$ commutes with the total Hamiltonian $\hat{H} = \hat{H}_{0} + \hat{V}$.
From this deduce the eigenstates and eigenvalues of $\hat{H}$.
In these eigenstates is the electron localized?

\textbf{Hint}: The $N$ distinct complex roots of $z^{N} = 1$ are given by $z_{n} = e^{i(2\pi n/N)}$ for $n=1,2,\ldots,N$, and the following identity holds:
\begin{eqnarray}
    \sum_{n=0}^{N} z^{n} = \frac{1 - z^{N+1}}{1 - z}
\end{eqnarray}
for complex $z$.

}

\sol{

(a) The eigen-equation for $\hat{R}$ is $\hat{R} \ket{\psi} = r \ket{\psi}$.
It is clear from the definition of $\hat{R}$ that $\hat{R}^{6} = \hat{\id}$, meaning that
\begin{eqnarray}
    \hat{R}^{6} \ket{\psi} = \hat{\id} \ket{\psi} = r^{6} \ket{\psi} \Rightarrow r^{6} = 1
.\end{eqnarray}
The eigenvalues of $\hat{R}$ are then just the sixth roots of unity: $\{ e^{i\pi/3}, e^{i 2\pi/3}, e^{i \pi}, e^{i 4\pi/3}, e^{i 5\pi/3}, 1 \}$.
If we denote the $n^{\rm th}$ sixth root of unity as $\omega_{n}$ for $n=1,\ldots,6$, the eigenvectors can be found from
\begin{eqnarray}
    \begin{pmatrix}
        -\omega_{n} & 1 & 0 & 0 & 0 & 0 \\
        0 & -\omega_n & 1 & 0 & 0 & 0 \\
        0 & 0 & -\omega_{n} & 1 & 0 & 0 \\
        0 & 0 & 0 & -\omega_{n} & 1 & 0 \\
        0 & 0 & 0 & 0 & -\omega_{n} & 1 \\
        1 & 0 & 0 & 0 & 0 & -\omega_n
    \end{pmatrix} 
    \begin{pmatrix}
        a_1 \\ a_2 \\ a_3 \\ a_4 \\ a_5 \\ a_6
    \end{pmatrix} 
    =
    0
.\end{eqnarray}
Equivalently, we have the system of equations
\begin{eqnarray}
    \begin{cases}
    -\omega_{n} a_1 + a_2 = 0 \\
    -\omega_{n} a_2 + a_3 = 0 \\
    -\omega_{n} a_3 + a_4 = 0 \\
    -\omega_{n}a_4 + a_5 = 0 \\
    -\omega_{n} a_5 + a_6 = 0 \\
    \phantom{+}a_1 -\omega_{n} a_6 = 0
    .\end{cases}
\end{eqnarray}
While it may not be clear from the system above, we know that these equations are linearly dependent with one free undetermined parameter since each eigenvalue has algebraic multiplicity $1$.
Thus, we can let $a_1$ be our undetermined parameter and solve for $a_2,\ldots,a_6$ in terms of $a_1$.
Denoting the eigenvector corresponding to eigenvalue $\omega_{n}$ as $\ket{\chi_{n}}$, we find
\begin{eqnarray}
    \chi_{n} = a_1 \begin{pmatrix}
        1\\
        \omega_{n} \\
        \omega_{n}^2 \\
        \omega_{n}^3 \\
        \omega_{n}^{4} \\
        \omega_{n}^{5} \\
    \end{pmatrix}    
.\end{eqnarray}
This is a very nice form for the eigenvectors.
Normalizing we find
\begin{eqnarray}
    \chi_{n}^{\dagger}\chi_{n} = 6 |a_1|^2 = 1 \Rightarrow a_1 = 1/\sqrt{6}
.\end{eqnarray}

(b) In the basis $\ket{\psi_1},\ldots,\ket{\psi_6}$
\begin{eqnarray}
    \hat{R} = \ket{\psi_1}\bra{\psi_2} + \ket{\psi_2}\bra{\psi_3} + \ket{\psi_3}\bra{\psi_4} + \ket{\psi_4}\bra{\psi_5} + \ket{\psi_5}\bra{\psi_6} + \ket{\psi_6}\bra{\psi_1} 
.\end{eqnarray}
It is clear then that
\begin{eqnarray}
    \hat{R}^{\dagger} = \ket{\psi_2}\bra{\psi_1} + \ket{\psi_3}\bra{\psi_2} + \ket{\psi_4}\bra{\psi_3} + \ket{\psi_5}\bra{\psi_4} + \ket{\psi_6}\bra{\psi_5} + \ket{\psi_1}\bra{\psi_6} 
.\end{eqnarray}
Thus, we have the actions
\begin{eqnarray}
    \hat{R}^{\dagger} \ket{\psi_1} = \ket{\psi_6}, \quad \hat{R}^{\dagger} \ket{\psi_2} = \ket{\psi_1}, \quad \ldots, \quad \hat{R}^{\dagger} \ket{\psi_6} = \ket{\psi_5}
.\end{eqnarray}

Furthermore, we have
\begin{eqnarray}
    R^{\dagger} R = 
   \begin{pmatrix}
        0 & 0 & 0 & 0 & 0 & 1 \\
        1 & 0 & 0 & 0 & 0 & 0 \\
        0 & 1 & 0 & 0 & 0 & 0 \\
        0 & 0 & 1 & 0 & 0 & 0 \\
        0 & 0 & 0 & 1 & 0 & 0 \\
        0 & 0 & 0 & 0 & 1 & 0
    \end{pmatrix}
    \begin{pmatrix}
        0 & 1 & 0 & 0 & 0 & 0 \\
        0 & 0 & 1 & 0 & 0 & 0 \\
        0 & 0 & 0 & 1 & 0 & 0 \\
        0 & 0 & 0 & 0 & 1 & 0 \\
        0 & 0 & 0 & 0 & 0 & 1 \\
        1 & 0 & 0 & 0 & 0 & 0 
    \end{pmatrix}
    = 
    \begin{pmatrix}
        1 & 0 & 0 & 0 & 0 & 0 \\
        0 & 1 & 0 & 0 & 0 & 0 \\
        0 & 0 & 1 & 0 & 0 & 0 \\
        0 & 0 & 0 & 1 & 0 & 0 \\
        0 & 0 & 0 & 0 & 1 & 0 \\
        0 & 0 & 0 & 0 & 0 & 1
    \end{pmatrix}
.\end{eqnarray}
Hence, $\hat{R}$ is unitary (in the matrix representation $R$ is orthogonal since it is real).

(c) It is clear that $[\hat{R},\hat{H}_0] = 0$ since in the eigenbasis of $\hat{H}_{0}$, $\hat{H}_{0} = E_0 \hat{\id}$.
Note that $\hat{V} = -a (\hat{R}^{\dagger} + \hat{R})$.
From this, it is clear that $[\hat{R},V] = 0$ since $\hat{R}$ commutes with itself and its adjoint\footnote{This is not generally true -- it is only the case here since $\hat{R}$ is unitary.}.
It is clear that $\{ \ket{\chi_{n}} \} $ forms a simultaneous orthonormal basis for $\hat{H}_0$ and $\hat{R}$.
From the fact that $\hat{V} = -a( \hat{R}^{\dagger} + \hat{R} )$ and that $\hat{R}^{\dagger} \ket{\chi_{n}} = \omega_{n}^{*} \ket{\chi_{n}}$, we have $\hat{V} \ket{\chi_{n}} = -(a/2) \Re(\omega_{n})$.
The eigenvalues of $\hat{H}$ are then given by $E_0 - (a/2)\cos(\pi n/3)$.

}

\end{document}
